---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  fig.path = "README/README-",
  fig.align = 'center',
  fig.retina=2, fig.width = 12, fig.height = 10.5,
  warning = FALSE,
  message = FALSE
)
options("width"=200)
```

iNEXT.4steps (R package)
=====

<h5 align="right">Latest version: `r Sys.Date()`</h5>

<font color="394CAE"><h3 color= 394CAE style = "font-weight: bold"> Introduction to iNEXT.4steps (R package): Excerpt from iNEXT.4steps User’s Guide </h3> </font>
<br>
<h5><b>Hsieh, Anne Chao, Kai-Hsiang Hu</b>
<br><br>
<i>Institute of Statistics, National Tsing Hua University, Hsin-Chu, Taiwan 30043</i>  
</h5>
<br>
iNEXT.4steps (iNterpolation and EXTrapolation for four steps) is an original R package available in [Github](https://github.com/AnneChao) which provide an easy complete biological analysis computation. In Chao, A., Y. Kubota, D. Zelený, C.-H. Chiu(2020), they proposed a complete biological analysis process: 

* `STEP1`. Sample completeness profiles. 

* `STEP2`. <font color=#FF6781> Size-based rarefaction and extrapolation analysis and the asymptotic diversity profile. </font>

* `STEP3`. Non-asymptotic coverage-based rarefaction and extrapolation analysis. 

* `STEP4`. Evenness among species abundances. 

These are the foundation of `iNEXT.4steps`. Here we will introduce functions about `STEP1` and `STEP4`, particularly. If you want to grasp `STEP2`, `STEP3`, then the related paper T.C. Hsieh, K. H. Ma, and Chao, A. (2016) or the packages (`iNEXT.3D`) will be the best choice to you. An online version of [iNEXT.4steps Online](https://chao.shinyapps.io/iNEXT_4steps/) is also available for users without an R background. Detailed information about all functions in iNEXT.4steps is provided in the iNEXT.4steps Manual in [iNEXT.4steps_vignettes](http://chao.stat.nthu.edu.tw/wordpress/wp-content/uploads/software/A%20Quick%20Introduction%20to%20iNEXT.4steps%20via%20Examples.html),which is available from [Anne Chao's website](http://chao.stat.nthu.edu.tw/wordpress/software_download/).

`iNEXT.3D` contains two major parts:

1. Non-asymptotic diversity:
- Sample-size-based <font color=#FF6781>(or size-based)</font> R/E sampling curves: `iNEXT3D` computes rarefied and extrapolated <font color=#FF6781>Taxonomic, Phylogenetic or Functional</font> diversity estimates under a specified sample size. This type of diversity curve plots the estimated diversity with respect to sample size.     

- Coverage-based R/E sampling curves: `iNEXT3D` computes rarefied and extrapolated <font color=#FF6781>Taxonomic, Phylogenetic or Functional</font> diversity estimates under a specified sample coverage. This type of diversity curve plots the estimated diversity with respect to sample coverage. 

2. Asymptotic diversity: asymptotic <font color=#FF6781>Taxonomic, Phylogenetic or Functional</font> diversity estimate with q-profile, time-profile, and tau-profile.


In `iNEXT.4steps` package, we provide other four functions for users to calculate <font color=#FF6781>and visualize the related biological statistics: `Completeness` and `Evenness`, for calculation. `ggCompleteness` and `ggEvenness`, for visualization .</font> The most comprehensive function `iNEXT4steps` gathers `iNEXT3D`, `AO3D`, `Completeness` and `Evenness` into a figure. In this document, we will give a quick introduction demonstrating how to run these functions. Detailed information about these function settings is provided in the `iNEXT.4steps` Manual. The theoretical basis about these biological statistics can be obtained from the following inference:
(http://chao.stat.nthu.edu.tw/wordpress/paper/135.pdf)


### SOFTWARE NEEDED TO RUN iNEXT.4steps IN R

-   Required: [R](http://cran.rstudio.com/)
-   Suggested: [RStudio IDE](http://www.rstudio.com/ide/download/)

### HOW TO RUN INEXT.4STEPS:

The `iNEXT.4steps` package can be downloaded from Anne Chao's [iNEXT.4steps_github](https://github.com/AnneChao/iNEXT.4steps) using the following commands. For a first-time installation, an additional visualization extension package (`ggplot2`) and (`iNEXT.3D`) from Anne Chao's github must be installed and loaded. 

```{r install, eval=FALSE}
## install the latest version from github
install.packages('devtools')
library(devtools)

install_github('AnneChao/iNEXT.4steps')
## import packages
library(iNEXT.4steps)

# install_github('AnneChao/iNEXT.3D')
# library(iNEXT.3D)

library(ggplot2)
```


## SIMPLE STEPS BRIEF

`Sample Completeness` represent the proportion of observed species in the population. Usually, the sampling data represents the abundant species in the population so that we will ignore the rare species. Here we will use Turing’s sample coverage theory to reconstruct the population proportion. Besides, sample completeness can correspond to order q, which is an weight index. When order q tends to zero, then we will give more weight to rare species. If order q tends to unity, then we will equally treat each species, which is also called sample coverage at unity. In contrast, if order q tends to larger than unity, we will give more weights to abundant species. By sample completeness, we can easily plot the estimated curve with respect to order q and associated 95% confidence interval.


`Asymptotic Diversity (or called Hill numbers)` is a statistic which is used to represent the biological diversity. Its direct meaning is to transform the non-homogeneous distribution into homogeneous distribution (species diversity). With this quantification criterian, we can easily analysis several communities for different data sources in an objective measure. It usually matches order q to give different weights so that we can focus on rare species or abundant species only. When we use observed data to calculate empirical diversity, we usually get an underestimated value. In our package, we will provide an estimated statistic from Anne Chao (2015), which can imitate a real population accurately. In addition, it can be related to order q index and associated 95% confidence interval.


`Interpolation and Extrapolation (iNEXT)` focuses on three measures of order q: species richness (q = 0), Shannon diversity (q = 1, the exponential of Shannon entropy), and Simpson diversity (q = 2, the inverse of Simpson concentration). For each diversity measures, iNEXT uses observed sample to compute expected diversity estimates and associated 95% confidence intervals according two different unit types of rarefaction and extrapolation (R/E):

1.	Sample-size-based R/E sampling curves versus diversity in each order q.

2.	Coverage-based R/E sampling curves versus diversity in each order q. 

For more particular usage about iNEXT, you can refer to T.C. Hsieh, K. H. Ma, and Chao, A. (2016). We won’t introduce details of iNEXT latter.


`Evenness` is an function to calculate whether a assemblage is uniform or not. We have sorted five main classes according to different transformation by species and diversity. In these five classes, they all have range from zero to one. When the value is close to zero, it means that the assemblage tends to uneven. On the contrary, when the value is close one, it means that the assemblage tends to even. Evenness can consider different order q under each classes. When order q tends to zero, we will focus on rare species. In other sides, when order q tends to far from zero, then we will do more weights on abundant species. If we use observed sampling data to calculate Evenness, we usually have a biased value because of unobserved rare species. But if we try to use asymptotic diversity for calculating Evenness, then we may also get a biased value when order q is near zero. Here, we propose a “standardized coverage” (named `Cmax`) as a judged criterion. `Cmax` means that we computes the diversity estimates for the minimum sample coverage among all samples extrapolated to double reference sizes.
Under Cmax, we can guarantee the accuracy of Evenness. According to this criterion, we can plot the Evenness curves versus order q and associated 95% confidence interval.



## DATA FORMAT/INFORMATION
Three types of data are supported:

1. Individual-based abundance data (`datatype="abundance"`): Input data for each assemblage/site include species abundances in an empirical sample of n individuals (“reference sample”). When there are N assemblages, input data consist of an S by N abundance matrix, or N lists of species abundances.  

2. Sampling-unit-based incidence data: There are two kinds of input data.  
 (2a) Incidence-raw data (`datatype="incidence_raw"`): for each assemblage, input data for a reference sample consisting of a species-by-sampling-unit matrix; each element in the raw matrix is 1 for a detection, and 0 otherwise. When there are N assemblages, input data consist of N lists of raw matrices, and each matrix is a species-by-sampling-unit matrix.  
 (2b) Incidence-frequency data (`datatype="incidence_freq"`): input data for each assemblage consist of species sample incidence frequencies (i.e., row sums of the corresponding incidence raw matrix). When there are N assemblages, input data consist of an (S+1) by N matrix, or N lists of species incidence frequencies. The first entry of each column/list must be the total number of sampling units, followed by the species incidence frequencies.


## MAIN FUNCTION: iNEXT4steps()

We first describe the main function `iNEXT4steps()` with default arguments: 

<br>
iNEXT4steps(data, diversity = "TD", q = seq(0, 2, 0.2), datatype = "abundance",
            nboot = 50, nT = NULL,
            PDtree = NULL, PDreftime = NULL, PDtype = "meanPD",
            FDdistM = NULL, FDtype = "AUC", FDtau = NULL,
            details = FALSE
)
<br>

The arguments of this function are briefly described below, and will explain details by illustrative examples in later text. 

```{r, echo=FALSE,warning=FALSE}
Des <- c("data"," (a) For `datatype = 'abundance'`, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. (b) For `datatype = 'incidence_freq'`, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. (c) For `datatype = 'incidence_raw'`, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. ",
"diversity"," selection of diversity type: 'TD' = Taxonomic diversity, 'PD' = Phylogenetic diversity, and 'FD' = Functional diversity.",
"q","	a numerical vector specifying the diversity orders. Default is c(0, 1, 2).",
"datatype"	,"data type of input data: individual-based abundance data (datatype = 'abundance'), sampling-unit-based incidence frequencies data (datatype = 'incidence_freq'), or species by sampling-units incidence matrix (datatype = 'incidence_raw') with all entries being 0 (non-detection) or 1 (detection).", 
"nboot"," a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.",
"nT","(required only when datatype = 'incidence_raw' and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as 'assemblage1', 'assemblage2',..., etc.",
"PDtree","	(required only when diversity = 'PD'), a phylogenetic tree in Newick format for all observed species in the pooled assemblage.",
"PDreftime", "(required only when diversity = 'PD'), a vector of numerical values specifying reference times for PD. Default is NULL (i.e., the age of the root of PDtree).",
"PDtype"," (required only when diversity = 'PD'), select PD type: PDtype = 'PD' (effective total branch length) or PDtype = 'meanPD' (effective number of equally divergent lineages). Default is 'meanPD', where meanPD = PD/tree depth.",
"FDdistM"," (required only when diversity = 'FD'), select FD type: FDtype = 'tau_values' for FD under specified threshold values, or FDtype = 'AUC' (area under the curve of tau-profile) for an overall FD which integrates all threshold values between zero and one. Default is 'AUC'.",
"FDtype"," (required only when diversity = 'FD'), select FD type: FDtype = 'tau_values' for FD under specified threshold values, or FDtype = 'AUC' (area under the curve of tau-profile) for an overall FD which integrates all threshold values between zero and one. Default is 'AUC'.",
"FDtau"," (required only when diversity = 'FD' and FDtype = 'tau_values'), a numerical vector between 0 and 1 specifying tau values (threshold levels). If NULL (default), then threshold is set to be the mean distance between any two individuals randomly selected from the pooled assemblage (i.e., quadratic entropy).",
"details"," a logical variable to decide whether do you want to print out the detailed value for each plots, default is `FALSE`.")
output <- 
  matrix(Des, 
         ncol=2, byrow = TRUE)

library(htmlTable)
htmlTable(output,
          header =  c("Argument","Description"),align = "l"
          )
```

Here data can be any diversity type, such as a list gathered by each vector, a data frame with each column versus a assemblage, a vector correspond to a assemblage. Besides, `datatype` means that you should be certain what type your data is.
Diversity dimension can be classified into three attributes diversity: 'Taxonomic diversity', 'Phylogenetic diversity', and 'Functional diversity'. So you should choose one `diversity` : `TD` means 'Taxonomic diversity', `PD` means 'Phylogenetic diversity' under a certain reference time(root), and `FD` means 'Functional diversity'. For the `”Functional diversity”`, `FDtype = "AUC"` is a conservative computation for functional diversity which integrate threshold between zero and one. And `FDtype = "tau_values"` computes functional diversity under certain threshold.
Besides, when you choose `diversity = “PD”`, you should input `PDtree` data. 
When you choose `'FD'`, you should also input `FDdistM` data, which represents distance matrix.

`datatype` has been introduced before. When the user use `datatype = "incidence_raw"` and class of data is 'data.frame' or 'matrix', user should input `nT` for each assemblage which represents sampling units.
`nboot` are applied to confidence interval, which is estimated by re-sampling method: bootstrap with `nboot` times.
Finally, `details` means a logical setting whether do you want to print out total information of five figures.

The output of `iNEXT4steps` will have three parts (if you sets `details = TRUE`): `$summary`, `$figure`, and `$details`. It may take you much time to calculate five figures if your data is large or you set a large number of `nboot.`


## Taxonomic Diversity ##

## Abundance-based ##

`“abundance Data”` usually indicated a random sampling source, and the individuals spread uniformly all over the area. If the species in your data has gathering effect, such as a kind of plant usually live near each other instead of spreading uniformly, then we will propose `datatype = “incidence_freq”` or `datatype = “incidence_raw”`. So user can follow a rule: `datatype = “abundance”` for animals and `datatype = “incidence”` for plants.
First, we try to calculate taxonomic diversity for `Spider`

### Spider abundance-based data ###

These data were sampled in a mountain forest ecosystem in the Bavarian Forest National Park, Germany (Thorn et al. 2016, 2017).
A total of 12 experimental plots were established in "closed forest" stands (6 plots) and "open forest" stands with naturally occurring gaps and edges (6 plots) to assess the effects of microclimate on communities of epigeal (ground-dwelling) spiders.
Epigeal spiders were sampled over three years with four pitfall traps in each plot, yielding a total of 3171 individuals belonging to 85 species recorded in the pooled habitat.

Data Source : A mountain forest ecosystem in the Bavarian Forest National Park, Germany (Thorn et al. 2016, 2017).


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, warning=FALSE)
library(iNEXT.4steps)
library(ggplot2)
```

```{r}
data(Spider)
out1 <- iNEXT4steps(data = Spider, diversity = "TD", datatype = "abundance")
out1$summary
out1$figure[[6]]
```

`$summary` lists all biological index according to Chao, A ., Y. Kubota, D. Zelený, C.-H. Chiu (2020). There are four parts corresponding to each step in the paper. They analysis and explain biological data from different and superimposed side. So users can easily compare difference between each assemblages. 

`$figure` visualize the statistics by continuous curves. From the above five figures, we provide a standard analysis process from figure (a) to figure (e). Except that figure (c) contains empirical statistics, all the other figures are estimated statistics. You can understand these estimated statistics through reference paper in the final page. 

`$details` contains four parts: `Sample Completeness`, `iNEXT`, `Asymptotic Diversity`, `Evenness`. \cr
Each meaning will be introduced in the individual function latter (except figure (b) and fugure (d)). 


## Incidence-based ##
Incidence data usually match by incidence-sampling-units. We split an area into several the same-size area and only care whether the species is detected or non-detected. According to this sampling method, `incidence_raw` data has value “zero” and “one”. And `incidence_freq` data is the total frequency incidence in each assemblage. You should input your total sampling units (plots number) in the first row. \cr
*Remark:* The phylogenetic diversity can only select `datatype = “incidence_raw”` for incidence-based data.

## Woody plants incidence-based data ##

The datasets considered here are a subset of The National Vegetation Database of Taiwan (AS-TW-001), sampled between 2003 and 2007 within the first national vegetation inventory project (Chiou et al. 2009).
Over 3600 vegetation plots, each 20x20-m in area, were set up in various locations in Taiwan, and all woody plant individuals taller than 2 meters were recorded in each plot.
For illustration here, we selected only plots belonging to two vegetation types (according to Li et al. 2013): Pyrenaria-Machilus subtropical winter monsoon forest and Chamaecyparis montane mixed cloud forest, sampled in the northern part of Taiwan (in ecoregions 7 and 8 according to Su 1985).

Data Source : The National Vegetation Database of Taiwan (AS-TW-001).

```{r align="center", out.width="100%"}
data(woody_plants)
out2 <- iNEXT4steps(data = woody_plants[,c(1,4)], diversity = "TD", datatype = "incidence_freq")
out2$summary
out2$figure[[6]]
```


## Phylogenetic Diversity ##
Next, we want to calculate `Phylogenetic diversity`. Here we provide `“abundance”` data : `“brazil”` as our example. `“brazil”` data has two main communities: "Edge", "Interior". \cr
Here we provide phylogenetic tree data and pair-wise distance matrix of `“brazil”` data. User can match these to fit phylogenetic diversity or functional diversity. \cr
Phylogenetic diversity needs a tree data which has newick-format, which is related to evolution history. We can choose a historic time `PDreftime` as beginning of evolution. Then phylogenetic diversity will be calculated in a time range from now to `PDreftime`

If user don’t input tree data, the error will jump out. So if user don’t have tree data, taxonomic diversity will be your best choice.

According to following R code, we can get similar output with taxonomic diversity.

If you want to use incidence-based data, you should use `datatype = "incidence_raw"` with incidence raw data.

```{r align="center", out.width="100%"}
data(brazil)
data(brazil_tree)
out3 <- iNEXT4steps(data = brazil, diversity = "PD", datatype = "abundance", nboot = 20, PDtree = brazil_tree)
out3$summary
out3$figure[[6]]
```


## Functional Diversity ##
Next, we use distance matrix of `brazil` data to calculate functional diversity. If user doesn’t input distance matrix, than the error will jump out. Distance matrix is a species pairwise matrix which means the distance between any two species. The value in distance matrix has range from zero to one. The higher the value is, the more different the species is. If the value is “0”, it means the same species in the criterion. (“0” appears in the diagonal term)

So, without distance matrix, taxonomic diversity will be you best choose. 
User can select `FDtau` to decide a threshold which can express the difference distance in any two species. If `FDtau` is small, then all species will be seen as different species. If `FDtau` is large, then all species will be seen as the same species.

```{r align="center", out.width="100%"}
data(brazil)
data(brazil_distM)
out4 <- iNEXT4steps(data = brazil, diversity = "FD", datatype = "abundance", nboot = 20, FDdistM = brazil_distM, FDtype = 'tau_values')
out4$summary
out4$figure[[6]]
```


### MAIN FUNCTION: Completeness()

Function `Completeness()` provides a easy way to compute estimated sample completeness with order q. It has default arguments: 

<br>
Completeness(data, q = seq(0, 2, 0.2), datatype = "abundance", nboot = 50, conf = 0.95, nT = NULL)
<br>

The function AO3D() can compute three type diversity(‘TD’,‘PD’,‘FD’),which including empirical diversity and asymptotic diversity. For either abundance data (datatype="abundance") or incidence data(datatype="incidence_freq" or "incidence_raw")with any specified level of q can be compute.

For example, the following command returns an empirical taxonomic diversity(‘TD’) and asymptotic taxonomic diversity(‘TD’) for dunes data, along with its confidence interval, for a specified q level from 0 to 2.

### GRAPHIC DISPLAYS FUNCTION: ggCompleteness()

We also provides a realized function `ggCompleteness` to plot the output from `Completeness()`:

<br>
ggCompleteness(output)
<br>

There are two simple example for funciton `Completeness` and function `ggCompleteness`, one for abundance-based and the other for incidence-based data. 

## Abundance-based ##

We use data Spider to calculate abundance-based sample completeness and plot it.

```{r align="center", out.width="100%"}
data(Spider)
out1 <- Completeness(data = Spider, datatype = "abundance")
ggCompleteness(out1)
```

## Incidence-based ##

We use data woody to calculate incidence-based sample completeness and plot it.


```{r align="center", out.width="100%"}
data(woody_plants)
out2 <- Completeness(data = woody_plants[,c(1,4)], datatype = "incidence_freq")
ggCompleteness(out2)
```


### MAIN FUNCTION: Evenness()

We describe the main function `Evenness()` with default arguments: 

<br>
Evenness(data, q = seq(0, 2, 0.2), datatype = "abundance", method = "Estimated",
         nboot = 50, conf = 0.95, nT = NULL, E.class = 1:5, C = NULL)
<br>

### GRAPHIC DISPLAYS FUNCTION: ggCompleteness()

We provide a function `ggEvenness()` to plot the output from `Evenness()`.

<br>
ggEvenness(output)
<br>

There are two simple example for funciton `Evenness` and function `ggEvenness`, one for abundance-based and the other for incidence-based data. 

## Abundance-based ##

We use data Spider to calculate abundance-based evenness and plot it.

```{r out.width="100%",fig.height=8}
data(Spider)
out1 <- Evenness(data = Spider, datatype = "abundance")
ggEvenness(out1)
```


## Incidence-based ##

We use data woody to calculate incidence-based evenness and plot it.


```{r out.width="100%",fig.height=8}
data(woody_plants)
out2 <- Evenness(data = woody_plants[,c(1,4)], datatype = "incidence_freq")
ggEvenness(out2)
```



### How to cite
- Chao, A., Y. Kubota, D. Zelený, C.-H. Chiu, C.-F. Li, B. Kusumoto, M. Yasuhara, S. Thorn, C.-L. Wei, M. J. Costello, and R. K. Colwell (2020). Quantifying sample completeness and comparing diversities among assemblages. Ecological Research, 35, 292-314.

### Referance

- `Sample Completeness`: Chao, A. and Jost. L. (2012) Coverage-based rarefaction and extrapolation: standardizing samples by completeness rather than size. Ecology, 93, 2533-2547.

- `Asymptotic Diversity`: Chao, A. and Jost, L. (2015). Estimating diversity and entropy profiles via discovery rates of new species. Methods in Ecology and Evolution, 6, 873-882. 

- `Interpolation and Extrapolation`: Chao, A., Gotelli, N. G., Hsieh, T. C., Sander, E. L., Ma, K. H., Colwell, R. K. and Ellison, A. M. (2014). Rarefaction and extrapolation with Hill numbers: a framework for sampling and estimation in species biodiversity studies. Ecological Monographs 84, 45-67. 

- `Interpolation and Extrapolation`: Chao, A., Henderson, P. A., Chiu, C.-H., Moyes, F., Hu, K.-H., Dornelas, M and Magurran, A. E. (2021). Measuring temporal change in alpha diversity: a framework integrating taxonomic, phylogenetic and functional diversity and the iNEXT.3D standardization. Methods in Ecology and Evolution, 12, 1926-1940.

- `Evenness`: Chao, A. and Ricotta, C. (2019). Quantifying evenness and linking it to diversity, beta diversity, and similarity. Ecology, 100(12), e02852.

- `iNEXT 4 steps`: Chao, A., Y. Kubota, D. Zelený, C.-H. Chiu, C.-F. Li, B. Kusumoto, M. Yasuhara, S. Thorn, C.-L. Wei, M. J. Costello, and R. K. Colwell (2020). Quantifying sample completeness and comparing diversities among assemblages. Ecological Research, 35, 292-314.



