---
title: "A Quick Introduction to iNEXT4steps via Examples"
author: "K. H. Hu, and Anne Chao"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    vignette: >
      %\VignetteIndexEntry{A Quick Introduction to iNEXT4steps via Examples}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina = 2,
                      fig.align = 'center',
                      fig.width = 12, fig.height = 8,
                      warning = FALSE, message = FALSE)
options("width" = 200)
if("ggthemes" %in% rownames(installed.packages()) == FALSE) {install.packages("ggthemes")}
if("gridExtra" %in% rownames(installed.packages()) == FALSE) {install.packages("gridExtra")}
devtools::load_all(".")
#library(iNEXT.3D)
library(ggplot2)
data("Spider")
data("brazil")
data("brazil_tree")
data("brazil_distM")
```

<font color=#FF6600>
</font>


`iNEXT4steps` (iNterpolation and EXTrapolation four steps) is an original R package which  realized the biological process completely. In Chao, A., Y. Kubota, D. Zelený, C.-H. Chiu(2020),
they proposed a complete biological analysis process:

`STEP1`. Sample completeness profiles
`STEP2`. Asymptotic analysis
`STEP3`. Non-asymptotic coverage-based rarefaction and extrapolation analysis
`STEP4`. Evenness among species abundances.


That’s the foundation of `iNEXT4steps` Here we will introduce functions about `STEP1`, `STEP4`, particularly. If you want to grasp `STEP1`, `STEP3`, then the related paper T.C. Hsieh, K. H. Ma, and Chao, A. (2016) will be the best choice for you.

1. Sample‐size‐based R/E sampling curves: `iNEXT` computes diversity estimates for rarefied and extrapolated samples up to an appropriate size. This type of sampling curve plots the diversity estimates with respect to sample size.     
2. Coverage‐based R/E sampling curves: `iNEXT` computes diversity estimates for rarefied and extrapolated samples with sample completeness (as measured by sample coverage) up to an appropriate coverage. This type of sampling curve plots the diversity estimates with respect to sample coverage. 

In the package, there are several functions for users to calculate the related biological statistics: `Sample Completeness`, `Evenness`, and the most comprehensive function `iNEXT4steps`. In this document, we provide a quick introduction demonstrating how to run these functions. Detailed information about these functions is provided in the `iNEXT4steps` guidelines, also available in [CRAN]. The theoretical basis about these biological statistics can be obtained from the following connection:
(http://chao.stat.nthu.edu.tw/wordpress/publication-list/)

Some proposed theoretical references to each statistics:
`Sample Completeness`: No.95 Chao, A. and Jost. L. (2012) .
`Asymptotic Diversity`: No.109 Chao, A. and Jost, L. (2015).
`Interplation and Extrapolation`: No.102 Chao, A., Gotelli, N. G., Hsieh, T. C. (2014).
`Evenness`: No.133 Chao, A. and Ricotta, C. (2019).
`iNEXT 4 steps`: No. 135 Chao, A., Y. Kubota, D. Zelený, C.-H. Chiu.


## SOFTWARE NEEDED TO RUN INEXT IN R
- Required: [R](https://cran.r-project.org/)
- Suggested: [RStudio IDE](https://www.rstudio.com/products/RStudio/#Desktop)

## HOW TO RUN INEXT:
The `iNEXT4steps` package has not available on [CRAN]. But can be downloaded with a standard R installation procedure using the following commands. For a first‐time installation, an additional visualization extension package (`ggplot2`) must be loaded. 


```{r eval=FALSE}
## install iNEXT package from CRAN
# install.packages("iNEXT.4steps")

## install the latest version from github
install.packages('devtools')
library(devtools)
install_github('KaiHsiangHu/iNEXT.3D')
install_github('KaiHsiangHu/iNEXT4.steps')

## import packages
library(iNEXT.4steps)
library(ggplot2)
```

**Remark**: In order to install `devtools` package, you should update R to the latest version. Also, to get `install_github` to work, you should install the `httr` package. 

An online version of iNEXT4steps(https://kaihsiang-hu.shinyapps.io/iNEXT_4steps_Online/) is also available for users without an R background.



**Simple steps brief**
`Sample Completeness` represent the proportion of observed species in the population. Usually, the sampling data represents the abundant species in the population so that we will ignore the rare species. Here we will match Turing’s sample coverage to reconstruct the population percentage. Besides, sample completeness can correspond to order q, which is an weight index. When order q tends to zero, then we will give more weight to rare species. If order q tends to unity, then we will equally treat each species, which is also called sample coverage at unity. In contrast, if order q tends to larger than unity, we will give more weights to abundant species. By sample completeness, we can easily plot the estimated curve with respect to order q and associated 95% confidence interval.


`Asymptotic Diversity (or called Hill numbers)` is a statistic which is used to represent the biological diversity. Its direct sense is the quantification amount of a assemblage. With this quantification amount, we can easily analysis several communities from different data sources. It usually matches order q to give different weights so that we can focus on rare species or abundant species. When we use observed data to calculate asymptotic diversity, we usually will get an underestimated value. In our package, we will provide an estimated statistic from Anne Chao (2015), which can copy a real population accurately. In addition, it can be related to order q index and associated 95% confidence interval.


`Interplation and Extrapolation (iNEXT)` focuses on three measures of Hill numbers of order q: species richness (q = 0), Shannon diversity (q = 1, the exponential of Shannon entropy), and Simpson diversity (q = 2, the inverse of Simpson concentration). For each diversity measures, iNEXT uses observed sample to compute expected diversity estimates and associated 95% confidence intervals according two different unit types of rarefaction and extrapolation (R/E):
1.	Sample-size-based R/E sampling curves versus diversity in each order q.
2.	Coverage-based R/E sampling curves versus diversity in each order q.
For more particular usage about iNEXT, you can refer to T.C. Hsieh, K. H. Ma, and Chao, A. (2016). We won’t introduce details of iNEXT latter.


`Evenness` is an indicator used to represent whether a data is uniform or not. We have sorted five main classes according to different formulas. All of them are combination by species and diversity. In these five classes, they all have range from zero to one. When the value is close to zero, it means that the assemblage tends to uneven. On the contrary, when the value is close one, it means that the assemblage tends to even. Evenness can match different order q under each classes. When order q tends to zero, we will focus on rare species. In other sides, when order q tends to far from zero, then we will do more weights on abundant species. If we use observed sampling data to calculate Evenness, we usually have a biased value because of unobserved rare species. But if we try to use asymptotic diversity for calculating Evenness, then we may also get a biased value under order q near zero. Here, we propose a “standardized coverage” (named Cmax) as a judged crierion. Cmax means that we extrapolate the sampling data double reference sample size and choose the minimum corresponding coverage (q = 1) under these sample sizes.
Under Cmax, we can guarantee the accuracy of Evenness. According to this criterion, we can plot the curves versus order q and associated 95% confidence interval.



**DATA FORMAT/INFORMATION**
Before introducing functions, we define what the datatype may be classified.
There are mainly two types of sampling data are supposed :
`1.	Individual‐based abundance data (datatype="abundance"`): 
Input data for each assemblage/site include samples species abundances in an observed sample of n individuals (“reference sample”). When there are N assemblages, input data consist of an S by N abundance matrix, or N lists of species abundances.

`2. Sampling-unit-based incidence data: There are two kinds of input data`.
`(1)	Incidence‐frequency data (datatype = "incidence_freq")`:
Input data for each assemblage consist of species sampling incidence frequencies. When there are N assemblages, input data consist of an S by N matrix, or N lists of species incidence frequencies. The first entry of each column or each list must be the total number of sampling units, followed by the species incidence frequencies.
`(2)	Incidence‐raw data (datatype = "incidence_raw")`: for each assemblage, input data for a reference sample consist of a species‐by‐sampling‐unit matrix.
When you select diversity type is 'TD', 'FD' or 'AUC', you can use `datatype = 'incidence_freq'`. And when you select diversity type is 'PD', you can only use `datatype = 'incidence_raw'`.



## MAIN FUNCTION: iNEXT4steps()

We first describe the main function `iNEXT4steps()` with default arguments: 

```{r eval=FALSE}
iNEXT4steps(data, diversity = "TD", datatype = "abundance", nboot = 50, nT = NULL,
            PDtree = NULL, PDreftime = NULL, PDtype = 'meanPD', FDdistM = NULL, FDtype = 'AUC', FDtau = NULL,
            p_row = 2, p_col = 3, details = FALSE)
```

The arguments of this function are briefly described below, and will be explained in more details by illustrative examples in later text. This main function computes diversity estimates of order q, the sample coverage estimates and related statistics for K (if `knots=K`) evenly‐spaced knots (sample sizes) between size 1 and the `endpoint`, where the endpoint is described below. Each knot represents a particular sample size for which diversity estimates will be calculated. By default, endpoint = double the reference sample size (total sample size for abundance data; total sampling units for incidence data). For example, if `endpoint = 10`, `knot = 4`, diversity estimates will be computed for a sequence of samples with sizes (1, 4, 7, 10).  
<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left"(a) For \code{datatype = "abundance"}, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For \code{datatype = "incidence_freq"}, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For \code{datatype = "incidence_raw"}, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. 
</td>
</tr>
<tr class="even">
<td align="center"><code>diversity</code></td>
<td align="left">selection of diversity type: \code{'TD'} = Taxonomic diversity, \code{'PD'} = Phylogenetic diversity, and \code{'FD'} = Functional diversity.</td>
</tr>
<tr class="odd">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (\code{datatype = "abundance"}), sampling-unit-based incidence frequencies data (\code{datatype = "incidence_freq"}), or species by sampling-units incidence matrix (\code{datatype = "incidence_raw"}) with all entries being 0 (non-detection) or 1 (detection)</td>
</tr>
<tr class="even">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="odd">
<td align="center"><code>nT</code></td>
<td align="left">(required only when \code{datatype = "incidence_raw"} and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc.</td>
</tr>
<tr class="even">
<td align="center"><code>PDtree</code></td>
<td align="left">(required only when \code{diversity = "PD"}), a phylogenetic tree in Newick format for all observed species in the pooled assemblage.</td>
</tr>
<tr class="odd">
<td align="center"><code>PDreftime</code></td>
<td align="left">(required only when \code{diversity = "PD"}), a vector of numerical values specifying reference times for PD. Default is \code{NULL} (i.e., the age of the root of PDtree).</td>
</tr>
<tr class="even">
<td align="center"><code>PDtype</code></td>
<td align="left">(required only when \code{diversity = "PD"}), select PD type: \code{PDtype = "PD"} (effective total branch length) or \code{PDtype = "meanPD"} (effective number of equally divergent lineages). Default is \code{"meanPD"}, where \code{meanPD = PD/tree depth}.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDdistM</code></td>
<td align="left">(required only when \code{diversity = "FD"}), a species pairwise distance matrix for all species in the pooled assemblage.</td>
</tr>
<tr class="even">
<td align="center"><code>FDtype</code></td>
<td align="left">(required only when \code{diversity = "FD"}), select FD type: \code{FDtype = "tau_values"} for FD under specified threshold values, or \code{FDtype = "AUC"} (area under the curve of tau-profile) for an overall FD which integrates all threshold values between zero and one. Default is \code{"AUC"}.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDtau</code></td>
<td align="left">(required only when \code{diversity = "FD"} and \code{FDtype = "tau_values"}), a numerical vector between 0 and 1 specifying tau values (threshold levels). If \code{NULL} (default), then threshold is set to be the mean distance between any two individuals randomly selected from the pooled assemblage (i.e., quadratic entropy).</td>
</tr>
<tr class="even">
<td align="center"><code>p_row</code></td>
<td align="left">row amount for 4 steps figure, default = 2.</td>
</tr>
<tr class="odd">
<td align="center"><code>p_col</code></td>
<td align="left">column number for 4 steps figure, default = 3.</td>
</tr>
<tr class="even">
<td align="center"><code>details</code></td>
<td align="left">a logical variable to decide whether do you want to print out the detailed value for each plots, default is \code{FALSE}.</td>
</tr>

</tbody>
</table>


Here data can be any diversity type, such as a list gathered by each vector, a data frame with each column versus a assemblage, a vector correspond to a assemblage. Besides, `datatype` means that you should certain what type your data is.
Diversity dimension can be classified into three attributes diversity: `Taxonomic diversity`, `Phylogenetic diversity`, and `Functional diversity`. So you should choose one `diversity` : `“TD”` means `”Taxonomic diversity”`, `“PD”` means `”Phylogenetic diversity”` under a certain reference time(root), and `“FD”` means `”Functional diversity”` under a certain threshold(dmean). For the `”Functional diversity”`, `"AUC"` is a conservative computation for functional diversity which integrate threshold between zero and union.
Besides, when you choose `diversity = “PD”`, you should input `tree` data. 
When you choose `'FD'`, you should also input `distM` data, which represents distance matrix.

`datatype` has been introduced before. When the user use `diversity = "PD"` and `datatype = "incidence_raw"`, user should input `nT` for each assemblage which represents sampling units.
`nboot` are applied to confidence interval, which is estimated by re-sampling method: bootstrap with `nboot` times.
Finally, `details` means a logical setting whether do you want to print out total information of five figures.

The output of `iNEXT4steps` will have three parts (if you sets details = TRUE), `$summary`, `$figure`, and `$details`. It may spend you much time to calculate five figures if your data is large or you set too much times of `nboot.`


## Taxonomic Diversity ##
`“abundance Data”` usually indicated a random sampling source, and the species are uniformly distributed all over the area. If your data belongs to a gathering assemblage, such as a kind of plant usually near each other instead of uniformly distributed, then we will propose you use `datatype = “incidence_freq”` nor `datatype = “abundance”` So we can classify biology easily in following rule: `datatype = “abundance”` for animals and `datatype = “incidence”` for plants.
First, we try to calculate taxonomic diversity for `Spider`

## Abundance-based ##

```{r}
data(Spider)
out1 <- iNEXT4steps(data = Spider, diversity = "TD", datatype = "abundance")
out1$summary
out1$figure[[6]]
```

`$summary` list all biological index according to Chao, A ., Y. Kubota, D. Zelený, C.-H. Chiu (2020). There are four parts corresponding to each step in the paper. They analysis and explain biological data from different and superimposed side. So users can easily compare difference between each assemblages.
`$figure` visualize the statistics by continuous curves. From the above five figures, we provide a standard analysis process from figure (a) to figure (e). Except that figure (c) contains empirical statistics, all the other figures are estimated statistics. You can understand these estimated statistics through reference paper in the final page.
$details contains four parts: `Sample Completeness`, `iNEXT`, `Asymptotic Diversity`, `Evenness`.
Each meanings will be introduced in the individual function latter (except iNEXT). If you are interested in `iNEXT` output, T.C. Hsieh, K. H. Ma, and Chao, A. (2016) will be your best choice.


## Incidence-based ##
Incidence data usually match by incident sampling. We split an area into several the same-size small area and only care whether the species is observed or not. According to this sampling method, our incidence data has value “zero” and “one” only except first row. Be careful, you should input your total sampling units (plots) in the first row.
In the past, there are `“incidence_freq”` and `“incidence_raw”` two types for incidence data. `“incidence_freq”` belongs to taxonomic diversity and functional diversity. `“incidence_raw”` belongs to phylogenetic diversity.

```{r}
data(woody_incid)
out2 <- iNEXT4steps(data = woody_incid[,c(1,4)], diversity = "TD", datatype = "incidence_freq")
out2$summary
out2$figure[[6]]
```


## Phylogenetic Diversity ##
Next, we want to caluculate `Phylogenetic diversity`. Here we provide `“abundance”` data : `“brazil”` as our example. `“brazil”` data has two main community: "Edge", "Interior". 
Beside, we also provide phylogenetic tree data paired distance matrix of `“brazil”` data. User can match these to fit phylogenetic diversity or functional diversity.
Phylogenetic diversity needs phylogenetic structure data, which is related to evalution history. We can choose a historic time as end of evalution. Then phylogenetic diversity will adapt timely range from now to indicator.
If user don’t input tree data, the error will jumped out. So if user don’t have tree data, taxonomic diversity will be your best choice.
According following R code, we can get similar output with taxonomic diversity.

If you want to use incidence-based data, you should use `datatype = "incidence_raw"` with incidence raw data.

```{r}
data(brazil)
data(brazil_tree)
out3 <- iNEXT4steps(data = brazil, diversity = "PD", datatype = "abundance", nboot = 0, PDtree = tree)
out3$summary
out3$figure[[6]]
```


## Functional Diversity ##
Next, we use distance matrix of `brazil` data to calculate functional diversity. If user don’t input distance matrix, than the error will jump out. Distance matrix is a paired square matrix which means the distance between any two species. The value in distance matrix has range from zero to one. Much higher the value is, much different the species is. Opposite relation can be applied, too. Apparently, if the value is “0”, it means the same species in the criterian. (“0” appears in the diagonal term)
So, without distance matrix, taxonomic diversity will be you best choose.

If you want to use incidence-based data, you should use `datatype = "incidence_freq"` with incidence frequency data.

```{r}
data(brazil)
data(brazil_distM)
out4 <- iNEXT4steps(data = brazil, diversity = "FD", datatype = "abundance", nboot = 20, FDdistM = distM)
out4$summary
out4$figure[[6]]
```



## License
The `iNEXT.4steps` package is licensed under the GPLv3. To help refine `iNEXT.4steps`, your comments or feedbacks would be welcome (please send them to Anne Chao or report an issue on iNEXT.4steps github [reop](https://github.com/KaiHsiangHu/iNEXT.4steps)). 

## References
- Chao, A. and Jost. L. (2012) Coverage-based rarefaction and extrapolation: standardizing samples by completeness rather than size. Ecology, 93, 2533-2547.

- Chao, A. and Jost, L. (2015). Estimating diversity and entropy profiles via discovery rates of new species. Methods in Ecology and Evolution, 6, 873-882.

- Chao, A., Gotelli, N. G., Hsieh, T. C., Sander, E. L., Ma, K. H., Colwell, R. K. and Ellison, A. M. (2014). Rarefaction and extrapolation with Hill numbers: a framework for sampling and estimation in species biodiversity studies. Ecological Monographs 84, 45-67.

- Chao, A. and Ricotta, C. (2019). Quantifying evenness and linking it to diversity, beta diversity, and similarity. Ecology, 100(12), e02852.

- Chao, A., Y. Kubota, D. Zelený, C.-H. Chiu, C.-F. Li, B. Kusumoto, M. Yasuhara, S. Thorn, C.-L. Wei, M. J. Costello, and R. K. Colwell (2020). Quantifying sample completeness and comparing diversities among assemblages. Ecological Research, 35, 292-314.

- Chao, A., Chiu, C.-H., Villéger, S., Sun, I.-F., Thorn, S., Lin, Y.-C., Chiang, J. M. and Sherwin, W. B. (2019). An attribute-diversity approach to functional diversity, functional beta diversity, and related (dis)similarity measures. Ecological Monographs, 89, e01343. 10.1002/ecm.1343.

- Chao, A., Chiu, C.-H., and Jost, L. (2014). Unifying species diversity, phylogenetic diversity, functional diversity, and related similarity/differentiation measures through Hill numbers. Annual Reviews of Ecology, Evolution, and Systematics, 45, 297-324. 

- Chao, A., Chiu, C.-H., Hsieh, T. C., Davis, T., Nipperess, D., and Faith, D. (2015) Rarefaction and extrapolation of phylogenetic diversity. Methods in Ecology and Evolution, 6, 380-388.

- T.C. Hsieh, K. H. Ma, and Chao, A. (2016). iNEXT: An R package for rarefaction and extrapolation of species diversity (Hill numbers). Methods in Ecology and Evolution, 7, 1451-1456.

