---
title: "A Quick Introduction to iNEXT.4steps via Examples"
author: "K. H. Hu, and Anne Chao"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    vignette: >
      %\VignetteIndexEntry{A Quick Introduction to iNEXT.4steps via Examples}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina = 2,
                      fig.align = 'center',
                      fig.width = 12, fig.height = 10.5,
                      warning = FALSE, message = FALSE)
options("width" = 200)
if("ggthemes" %in% rownames(installed.packages()) == FALSE) {install.packages("ggthemes")}
if("gridExtra" %in% rownames(installed.packages()) == FALSE) {install.packages("gridExtra")}
#devtools::load_all(".")
library(iNEXT.4steps)
library(ggplot2)
data("Spider")
data("brazil")
data("brazil_tree")
data("brazil_distM")
```

<font color=#FF6600>
</font>


`iNEXT.4steps` (iNterpolation and EXTrapolation for four steps of biodiversity) is an original R package which provide an easy complete biological analysis computation. In Chao et al. (2020) paper, they proposed a complete biological analysis process:

* `STEP1`. Sample completeness profiles. 

* `STEP2`. Size-based rarefaction and extrapolation analysis and the asymptotic diversity profile. 

* `STEP3`. Non-asymptotic coverage-based rarefaction and extrapolation analysis. 

* `STEP4`. Evenness among species abundances. 


These are the foundation of `iNEXT.4steps`. Here we will introduce functions about `STEP1` and `STEP4`, particularly. If you want to grasp the functions of `STEP2`, `STEP3` individually, then please search the related paper Hsieh et al. (2016) or the package [iNEXT.3D](https://github.com/AnneChao/iNEXT.3D) from Anne Chao's github to get more details. `iNEXT.3D` contains two major parts:

1. Sample-size-based (or size-based) R/E sampling curves: `iNEXT3D` computes rarefied and extrapolated `taxonomic`, `phylogenetic`, or `functional` diversity estimates under a specified sample size. This type of diversity curve plots the estimated diversity with respect to sample size.     

2. Coverage-based R/E sampling curves: `iNEXT3D` computes rarefied and extrapolated `taxonomic`, `phylogenetic`, or `functional` diversity estimates under a specified sample coverage. This type of diversity curve plots the estimated diversity with respect to sample coverage. 


In `iNEXT.4steps` package, we provide other four functions for users to calculate and visualize the related biological statistics: `Completeness` and `Evenness` for calculation, and  `ggCompleteness` and `ggEvenness` for visualization. The most comprehensive function `iNEXT4steps` gathers `iNEXT3D`, `AO3D`, `Completeness` and `Evenness` into a summary table and a  figure. In this document, we will give a quick introduction demonstrating how to run these functions. Detailed information about these function settings is provided in the `iNEXT.4steps` manual, which will be submitted to [CRAN]. The theoretical basis about these biological statistics can be obtained from the following inference:
(http://chao.stat.nthu.edu.tw/wordpress/paper/135.pdf)


## HOW TO CITE iNEXT
If you publish your work based on the results from the `iNEXT.4steps` package, you should make references to the following methodology paper (Chao et al. 2020):

- Chao, A., Y. Kubota, D. Zelený, C.-H. Chiu, C.-F. Li, B. Kusumoto, M. Yasuhara, S. Thorn, C.-L. Wei, M. J. Costello, and R. K. Colwell (2020). Quantifying sample completeness and comparing diversities among assemblages. Ecological Research, 35, 292-314.


## SOFTWARE NEEDED TO RUN INEXT IN R
- Required: [R](https://cran.r-project.org/)
- Suggested: [RStudio IDE](https://www.rstudio.com/products/RStudio/#Desktop)

## HOW TO RUN INEXT:
The `iNEXT.4steps` package will be submitted to [CRAN] and can be downloaded from Anne Chao's [iNEXT.4steps_github](https://github.com/AnneChao/iNEXT.4steps) using the following commands. For a first-time installation, a related R package (`iNEXT.3D`) and an additional visualization extension package (`ggplot2`) must be installed and loaded. 


```{r eval=FALSE}
# install_github('AnneChao/iNEXT.3D')
# library(iNEXT.3D)

## install iNEXT.4steps package from CRAN
# install.packages("iNEXT.4steps")  # coming soon

## install the latest version from github
install.packages('devtools')
library(devtools)

install_github('AnneChao/iNEXT.4steps')
## import packages
library(iNEXT.4steps)

library(ggplot2)
```

An online version of iNEXT.4steps(https://chao.shinyapps.io/iNEXT_4steps/) is also available for users without R background.

## SIMPLE STEPS BRIEF


Step 1: `Sample completeness profile` 

`Sample Completeness` represent the proportion of observed species in the population (Chao et al., 2020). Usually, the sampling data represents the abundant species in the population so that we will ignore the rare species. Here we will use Turing’s sample coverage theory to reconstruct the population proportion. Besides, sample completeness can correspond to order q, which is an weight index. When order q tends to zero, then we will give more weight to rare species. If order q tends to unity, then we will equally treat each species, which is also called sample coverage at unity. In contrast, if order q tends to larger than unity, we will give more weights to abundant species. By sample completeness, we can easily plot the estimated curve with respect to order q and associated 95% confidence interval.


Step 2.1 and step 3: `Size-based and coverage-based Interpolation and Extrapolation` 

`Interpolation and Extrapolation (iNEXT)` focuses on three measures of order q: species richness (q = 0), Shannon diversity (q = 1, the exponential of Shannon entropy), and Simpson diversity (q = 2, the inverse of Simpson concentration) (Chao and Jost, 2012; Chao et al. 2014). For each diversity measures, iNEXT uses observed sample to compute expected diversity estimates and associated 95% confidence intervals according two different unit types of rarefaction and extrapolation (R/E):

1.	Sample-size-based R/E sampling curves versus diversity in each order q.

2.	Coverage-based R/E sampling curves versus diversity in each order q. 

For more particular usage about iNEXT, please refer to Hsieh et al. (2016). We won’t introduce details of iNEXT latter.


Step 2.2: `Asymptotic diversity profile` 

`Asymptotic Diversity (or called Hill numbers)` is a statistic which is used to represent the biological diversity. Its direct meaning is to transform the non-homogeneous distribution into homogeneous distribution (effective number of species). With this quantification criterian, we can easily analysis several communities for different data sources in an objective measure. It usually matches order q to give different weights so that we can focus on rare species or abundant species. When we use observed data to calculate empirical diversity, we usually get an underestimated value. In our package, we will provide an estimated statistic from Chao and Jost (2015), which can imitate a real population accurately. In addition, it can be related to order q index and associated 95% confidence interval.


Step 4: `Evenness profile` 

`Evenness` is an function to calculate whether a assemblage is uniform or not. We have sorted five main classes according to different transformation by species and diversity (Chao and Ricotta, 2019). In these five classes, they all have range from zero to one. When the value is close to zero, it means that the assemblage tends to uneven. On the contrary, when the value is close one, it means that the assemblage tends to uniform. Evenness considers different order q under each classes. When order q tends to zero, we will focus on rare species. In other sides, when order q tends to far from zero, then we will give more weights on abundant species. If we use observed sampling data to calculate Evenness, we usually have a biased value because of unobserved rare species. But if we try to use asymptotic diversity for calculating Evenness, then we may get a upper bound value. Here, we propose a “standardized coverage” (named `Cmax`) as a judged criterion. `Cmax` means that we computes the diversity estimates for the minimum sample coverage among all samples extrapolated to double reference sizes.
Under Cmax, we can guarantee the accuracy of Evenness. According to this criterion, we can plot the Evenness curves versus order q and associated 95% confidence interval.


The steps 2.1, 2.2, and 3 can also be promoted to phylogenetic diversity and functional diversity under attribute diversity framework; see Chao et al. (2019, 2021) to get more details.


## DATA FORMAT/INFORMATION
Three types of data are supported:

1. Individual-based abundance data (`datatype="abundance"`): Input data for each assemblage/site include species abundances in an empirical sample of n individuals (“reference sample”). When there are N assemblages, input data consist of an S by N abundance matrix, or N lists of species abundances.  

2. Sampling-unit-based incidence data: There are two kinds of input data.  
  (a) Incidence-raw data (`datatype="incidence_raw"`): for each assemblage, input data for a reference sample consisting of a species-by-sampling-unit matrix; each element in the raw matrix is 1 for a detection, and 0 otherwise. When there are N assemblages, input data consist of N lists of raw matrices, and each matrix is a species-by-sampling-unit matrix.  
  (b) Incidence-frequency data (`datatype="incidence_freq"`): input data for each assemblage consist of species sample incidence frequencies (i.e., row sums of the corresponding incidence raw matrix). When there are N assemblages, input data consist of an (S+1) by N matrix, or N lists of species incidence frequencies. The first entry of each column/list must be the total number of sampling units, followed by the species incidence frequencies.


## MAIN FUNCTION: iNEXT4steps()

We first describe the main function `iNEXT4steps()` with default arguments: 

```{r eval=FALSE}
iNEXT4steps(data, diversity = "TD", q = seq(0, 2, 0.2), datatype = "abundance",
            nboot = 50, nT = NULL,
            PDtree = NULL, PDreftime = NULL, PDtype = "meanPD",
            FDdistM = NULL, FDtype = "AUC", FDtau = NULL,
            details = FALSE
)
```

The arguments of this function are briefly described below, and will explain details by illustrative examples in later text. 
<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For `datatype = "abundance"`, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For `datatype = "incidence_freq"`, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For `datatype = "incidence_raw"`, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. 
</td>
</tr>
<tr class="even">
<td align="center"><code>diversity</code></td>
<td align="left">selection of diversity type: `TD` = Taxonomic diversity, `PD` = Phylogenetic diversity, and `FD` = Functional diversity.</td>
</tr>
<tr class="odd">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders. Default is (0,0.2,0.4,...,2).</td>
</tr>
<tr class="even">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (`datatype = "abundance"`), sampling-unit-based incidence frequencies data (`datatype = "incidence_freq"`), or species by sampling-units incidence matrix (`datatype = "incidence_raw"`) with all entries being 0 (non-detection) or 1 (detection)</td>
</tr>
<tr class="odd">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="even">
<td align="center"><code>nT</code></td>
<td align="left">(required only when `datatype = "incidence_raw"` and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc.</td>
</tr>
<tr class="odd">
<td align="center"><code>PDtree</code></td>
<td align="left">(required only when `diversity = "PD"`), a phylogenetic tree in Newick format for all observed species in the pooled assemblage.</td>
</tr>
<tr class="even">
<td align="center"><code>PDreftime</code></td>
<td align="left">(required only when `diversity = "PD"`), a vector of numerical values specifying reference times for PD. Default is `NULL` (i.e., the age of the root of PDtree).</td>
</tr>
<tr class="odd">
<td align="center"><code>PDtype</code></td>
<td align="left">(required only when `diversity = "PD"`), select PD type: `PDtype = "PD"` (effective total branch length) or `PDtype = "meanPD"` (effective number of equally divergent lineages). Default is `meanPD`, where `meanPD = PD/tree depth`.</td>
</tr>
<tr class="even">
<td align="center"><code>FDdistM</code></td>
<td align="left">(required only when `diversity = "FD"`), a species pairwise distance matrix for all species in the pooled assemblage.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDtype</code></td>
<td align="left">(required only when `diversity = "FD"`), select FD type: `FDtype = "tau_values"` for FD under specified threshold values, or `FDtype = "AUC"` (area under the curve of tau-profile) for an overall FD which integrates all threshold values between zero and one. Default is `AUC`.</td>
</tr>
<tr class="even">
<td align="center"><code>FDtau</code></td>
<td align="left">(required only when `diversity = "FD"` and `FDtype = "tau_values"`), a numerical vector between 0 and 1 specifying tau values (threshold levels). If `NULL` (default), then threshold is set to be the mean distance between any two individuals randomly selected from the pooled assemblage (i.e., quadratic entropy).</td>
</tr>
<tr class="odd">
<td align="center"><code>details</code></td>
<td align="left">a logical variable to decide whether do you want to print out the detailed value for each plots, default is `FALSE`.</td>
</tr>

</tbody>
</table>


Here the input data format can be several formats, such as a vector gathered by factors (`abundance` and `incidence_freq`), a matrix/data frame with species versus a assemblage (`abundance` `incidence_freq`, and `incidence_raw`), a list of several vectors (`abundance` and `incidence_freq`), or a list correspond to a assemblage (`incidence_raw`). `data` should  comform the format of each datatype. When `datatype = "incidence_raw"` and class of data is matrix/data frame, user should input `nT` for each assemblage which represents sampling units.

`diversity`  contains three attributes diversity dimensions: `Taxonomic diversity`, `Phylogenetic diversity`, and `Functional diversity`. User should choose one `diversity`: `TD` means `Taxonomic diversity`, `PD` means `Phylogenetic diversity` under a specified reference time (default is root height), and `FD` means `Functional diversity`. For the `”Functional diversity”`, `FDtype = "AUC"` consider overall functional diversity which integrates all threshold values between 0 to 1. And `FDtype = "tau_values"` computes functional diversity under specified thresholds.

When `diversity = “PD”`, user should input `PDtree` Newick format data for all observed species. When `diversity = “FD”`, user should input `FDdistM` data matrix. Each element of the matrix is the pairwise distance between any two observed species. And the species identification names should be listed as row names and column names of the matrix.

`nboot` is applied to get confidence interval, which is estimated by bootstrap method. `details` means a logical setting whether print out the computation value of all figures.

The output of `iNEXT4steps` will have three parts (if `details = TRUE`): `$summary`, `$figure`, and `$details`. It may take some time to compute five figures when data size is large or `nboot` is large.


## Taxonomic Diversity ##
`“abundance Data”` is used for a random sampling scheme. If the species has aggregation effect, such as trees or plants, then set `datatype = “incidence_freq”` or `datatype = “incidence_raw”`. First, we use data `Spider` to compute taxonomic diversity.

## Abundance-based ##

Datasets `Spider` were sampled in a mountain forest ecosystem in the Bavarian Forest National Park, Germany (Thorn et al. 2016, 2017).
A total of 12 experimental plots were established in "closed forest" stands (6 plots) and "open forest" stands with naturally occurring gaps and edges (6 plots) to assess the effects of microclimate on communities of epigeal (ground-dwelling) spiders. Epigeal spiders were sampled over three years with four pitfall traps in each plot, yielding a total of 3171 individuals belonging to 85 species recorded in the pooled habitat. More details refer to data Source : A mountain forest ecosystem in the Bavarian Forest National Park, Germany (Thorn et al. 2016, 2017).



```{r align="center",out.width="100%"}
data(Spider)
out1 <- iNEXT4steps(data = Spider, diversity = "TD", datatype = "abundance")
out1$summary
out1$figure[[6]]
```

`$summary` lists all biological summaries according to Chao et al. (2020). There are four parts corresponding to each step in the paper. They analysis and explain biological data from different and superimposed side. User can easily compare difference between each assemblages. 

`$figure` visualize the statistics by continuous curves. From the above five figures, `iNEXT4stpes` provides a standard analysis process from figure (a) to figure (e). User can analyze the process of biodiversity through these figures. 

`$details` contains four parts: `Sample Completeness`, `iNEXT`, `Asymptotic Diversity`, `Evenness`. They are the computing values which are used to plot each figure in `$figure`. 


## Incidence-based ##

Incidence data is matched by incidence-sampling-units. We split a space into several quadrats and only record whether the species is detected or undetected in each quadrat. According to this sampling scheme, `incidence_raw` data has only value “zero” (undetected) or “one” (detected) in matrix/data frame (species by assemblages). `incidence_freq` data is the total incidence frequency for each species (i.e., row sums of the corresponding incidence raw matrix). `incidence_freq` data should contain total sampling units (number of quadrats) in the first row/entry.

*Remark:* The phylogenetic diversity can only select `datatype = “incidence_raw”` for incidence-based data.

Datasets `Woody plants` are a subset of The National Vegetation Database of Taiwan (AS-TW-001), sampled between 2003 and 2007 within the first national vegetation inventory project (Chiou et al. 2009). Over 3600 vegetation plots, each 20x20-m in area, were set up in various locations in Taiwan, and all woody plant individuals taller than 2 meters were recorded in each plot. For illustration here, we selected only plots belonging to two vegetation types (according to Li et al. 2013): Pyrenaria-Machilus subtropical winter monsoon forest and Chamaecyparis montane mixed cloud forest, sampled in the northern part of Taiwan (in ecoregions 7 and 8 according to Su 1985).


```{r align="center", out.width="100%"}
data(woody_plants)
out2 <- iNEXT4steps(data = woody_plants[,c(1,4)], diversity = "TD", datatype = "incidence_freq")
out2$summary
out2$figure[[6]]
```


## Phylogenetic Diversity ##
Here use abundance data : `“brazil”` as example. `“brazil”` data has two main communities: "Edge", "Interior". Here we also provide phylogenetic tree data and pairwise distance matrix of `“brazil”` data to compute phylogenetic diversity and functional diversity. 

According to following R code, user can get similar output with taxonomic diversity.


```{r align="center", out.width="100%"}
data(brazil)
data(brazil_tree)
out3 <- iNEXT4steps(data = brazil, diversity = "PD", datatype = "abundance", nboot = 20, PDtree = brazil_tree)
out3$summary
out3$figure[[6]]
```


## Functional Diversity ##
Here abundance data `brazil` and its pairwise distance matrix to compute functional diversity. Under `FDtype = "tau_values"`, user can key in `FDtau` as thresholds. If `FDtau` is small,  all species will tend to different functional groups. If `FDtau` is large, all species will tend to the same functional group.

```{r align="center", out.width="100%"}
data(brazil)
data(brazil_distM)
out4 <- iNEXT4steps(data = brazil, diversity = "FD", datatype = "abundance", nboot = 20, FDdistM = brazil_distM, FDtype = 'tau_values')
out4$summary
out4$figure[[6]]
```

## MAIN FUNCTION: Completeness()

`iNEXT.4steps` provides function `Completeness()` to compute estimated sample completeness with order q. The arguments is below: 

```{r eval=FALSE}
Completeness(data, q = seq(0, 2, 0.2), datatype = "abundance", nboot = 50, conf = 0.95, nT = NULL)
```

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For `datatype = "abundance"`, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For `datatype = "incidence_freq"`, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For `datatype = "incidence_raw"`, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. 
</td>
</tr>
<tr class="even">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders. Default is (0, 0.2, 0.4,..., 2).</td>
</tr>
<tr class="odd">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (`datatype = "abundance"`), sampling-unit-based incidence frequencies data (`datatype = "incidence_freq"`), or species by sampling-units incidence matrix (`datatype = "incidence_raw"`) with all entries being 0 (non-detection) or 1 (detection)</td>
</tr>
<tr class="even">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="odd">
<td align="center"><code>conf</code></td>
<td align="left">a positive number < 1 specifying the level of confidence interval. Default is 0.95.</td>
</tr>
<tr class="even">
<td align="center"><code>nT</code></td>
<td align="left">(required only when `datatype = "incidence_raw"` and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc.</td>
</tr>

</tbody>
</table>

## MAIN FUNCTION: ggCompleteness()

`iNEXT.4steps` also provides a visualized function `ggCompleteness` to plot the output from `Completeness()`:

```{r eval=FALSE}
ggCompleteness(output)
```

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>output</code></td>
<td align="left">a table generated from function `Completeness`.
</td>

</tbody>
</table>


There are two simple examples for functions `Completeness` and `ggCompleteness`. One is abundance-based and the other is incidence-based data. 

## Abundance-based ##

Use abundance data `Spider` to compute sample completeness and plot it.

```{r out.width="70%",fig.height=8}
data(Spider)
out1 <- Completeness(data = Spider, datatype = "abundance")
ggCompleteness(out1)
```


## Incidence-based ##

Use incidence frequency data `woody plants` to compute sample completeness and plot it.

```{r out.width="70%",fig.height=8}
data(woody_plants)
out2 <- Completeness(data = woody_plants[,c(1,4)], datatype = "incidence_freq")
ggCompleteness(out2)
```



## MAIN FUNCTION: Evenness()

`iNEXT.4steps` provides the function `Evenness()` to compute observed (empirical) eveness or estimated evenness under specified sample coverage. The arguments is below: 

```{r eval=FALSE}
Evenness(data, q = seq(0, 2, 0.2), datatype = "abundance", method = "Estimated",
         nboot = 50, conf = 0.95, nT = NULL, E.class = 1:5, C = NULL)
```

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For `datatype = "abundance"`, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For `datatype = "incidence_freq"`, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For `datatype = "incidence_raw"`, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. 
</td>
</tr>
<tr class="even">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders. Default is (0,0.2,0.4,...,2).</td>
</tr>
<tr class="odd">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (`datatype = "abundance"`), sampling-unit-based incidence frequencies data (`datatype = "incidence_freq"`), or species by sampling-units incidence matrix (`datatype = "incidence_raw"`) with all entries being 0 (non-detection) or 1 (detection)</td>
</tr>
<tr class="even">
<td align="center"><code>method</code></td>
<td align="left">a binary calculation method with 'Estimated' or 'Empirical'.</td>
</tr>
<tr class="odd">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="even">
<td align="center"><code>conf</code></td>
<td align="left">a positive number < 1 specifying the level of confidence interval. Default is 0.95.</td>
</tr>
<tr class="odd">
<td align="center"><code>nT</code></td>
<td align="left">(required only when `datatype = "incidence_raw"` and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc.</td>
</tr>
<tr class="even">
<td align="center"><code>E.class</code></td>
<td align="left">an integer vector between 1 to 5. There are five transformation for evenness in Chao and Ricotta (2019). Default is 1:5.</td>
</tr>
<tr class="odd">
<td align="center"><code>C</code></td>
<td align="left">(required only when `method = 'Estimated'`) a standardized coverage for calculating evenness. If `NULL`, then this function computes the diversity estimates for the minimum sample coverage among all samples extrapolated to double reference sizes (`Cmax`).</td>
</tr>

</tbody>
</table>

## MAIN FUNCTION: ggEvenness()

`iNEXT.4steps` provide a function `ggEvenness()` to plot the output from `Evenness()`.

```{r eval=FALSE}
ggEvenness(output)
```

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>output</code></td>
<td align="left">a table generated from function `Evenness`.
</td>

</tbody>
</table>

There are two simple examples for functions `Evenness` and `ggEvenness`. One is abundance-based and the other is incidence-based data. 

## Abundance-based ##

Use abundance data `Spider` to calculate estimated evenness under `C = Cmax` and plot it.

```{r out.width="100%",fig.height=8}
data(Spider)
out1 <- Evenness(data = Spider, datatype = "abundance")
ggEvenness(out1)
```


## Incidence-based ##

Use incidence frequency data `Woody plants` to calculate estimated evenness under `C = Cmax` and plot it.


```{r out.width="100%",fig.height=8}
data(woody_plants)
out2 <- Evenness(data = woody_plants[,c(1,4)], datatype = "incidence_freq")
ggEvenness(out2)
```



## License
The iNEXT.4steps package is licensed under the GPLv3. To help refine `iNEXT.4steps`, your comments or feedback would be welcome (please send them to Anne Chao or report an issue on the iNEXT.4steps github [iNEXT.4steps_github](https://github.com/AnneChao/iNEXT.4steps). 

## References
- Chao, A., Chiu, C.-H., Villéger, S., Sun, I.-F., Thorn, S., Lin, Y.-C., Chiang, J. M. and Sherwin, W. B. (2019). An attribute-diversity approach to functional diversity, functional beta diversity, and related (dis)similarity measures. Ecological Monographs, 89, e01343. 10.1002/ecm.1343.

- Chao, A., Gotelli, N. G., Hsieh, T. C., Sander, E. L., Ma, K. H., Colwell, R. K. and Ellison, A. M. (2014). Rarefaction and extrapolation with Hill numbers: a framework for sampling and
estimation in species biodiversity studies. Ecological Monographs 84, 45-67. 

- Chao, A., Henderson, P. A., Chiu, C.-H., Moyes, F., Hu, K.-H., Dornelas, M and Magurran, A. E. (2021). Measuring temporal change in alpha diversity: a framework integrating taxonomic, phylogenetic and functional diversity and the iNEXT.3D standardization. Methods in Ecology and Evolution, 12, 1926-1940.

- Chao, A. and Jost. L. (2012) Coverage-based rarefaction and extrapolation: standardizing samples by completeness rather than size. Ecology, 93, 2533-2547.

- Chao, A. and Jost, L. (2015). Estimating diversity and entropy profiles via discovery rates of new species. Methods in Ecology and Evolution, 6, 873-882.

- Chao, A. and Ricotta, C. (2019). Quantifying evenness and linking it to diversity, beta diversity, and similarity. Ecology, 100(12), e02852.

- Chao, A., Y. Kubota, D. Zelený, C.-H. Chiu, C.-F. Li, B. Kusumoto, M. Yasuhara, S. Thorn, C.-L. Wei, M. J. Costello, and R. K. Colwell (2020). Quantifying sample completeness and comparing diversities among assemblages. Ecological Research, 35, 292-314.

- Chiou, C.-R., Hsieh, C.-F., Wang, J.-C., Chen, M.-Y., Liu, H.-Y., Yeh, C.-L., … Song, M. G.-Z. (2009). The first national vegetation inventory in Taiwan. Taiwan Journal of Forest Science, 24, 295–302.

- Li, C.-F., Chytrý, M., Zelený, D., Chen, M. -Y., Chen, T.-Y., Chiou, C.-R., … Hsieh, C.-F. (2013). Classification of Taiwan forest vegetation. Applied Vegetation Science, 16, 698–719. https://doi.org/10.1111/avsc.12025

- Su, H. -J. (1985). Studies on the climate and vegetation types of the
natural forests in Taiwan (III) A Scheme of Geographical Climatic Regions. Quarterly Journal of Chinese Forestry, 18, 33–44.

- T.C. Hsieh, K. H. Ma, and Chao, A. (2016). iNEXT: An R package for rarefaction and extrapolation of species diversity (Hill numbers). Methods in Ecology and Evolution, 7, 1451-1456.

- Thorn, S., Bässler, C., Svoboda, M., & Müller, J. (2017). Effects of
natural disturbances and salvage logging on biodiversity - lessons from the bohemian Forest. Forest Ecology and Management, 388, 113–119. https://doi.org/10.1016/j.foreco.2016.06.006

- Thorn, S., Bußler, H., Fritze, M. -A., Goeder, P., Müller, J., Weiß, I., & Seibold, S. (2016). Canopy closure determines arthropod assemblages in microhabitats created by windstorms and salvage logging. Forest Ecology and Managemen


