---
title: "A Quick Introduction to iNEXT.4steps via Examples"
author: "Anne Chao and Kai-Hsiang Hu"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    vignette: >
      %\VignetteIndexEntry{A Quick Introduction to iNEXT.4steps via Examples}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina = 2,
                      fig.align = 'center',
                      fig.width = 12, fig.height = 10.5,
                      warning = FALSE, message = FALSE)
options("width" = 200, digits = 3)

library(iNEXT.4steps)
library(ggplot2)
data("Spider")
data("Brazil_abun_data")
data("Brazil_phylo_tree")
data("Brazil_distance_matrix")
```

<font color=#FF6600>
</font>


`iNEXT.4steps` (iNterpolation and EXTrapolation for four steps of biodiversity) is an original R package which provide an easy complete biological analysis computation. In Chao et al. (2020) paper, they proposed a complete biological analysis process:

* `STEP1`. Sample completeness profiles. 

* `STEP2`. Size-based rarefaction and extrapolation analysis and the asymptotic diversity profile. 

* `STEP3`. Non-asymptotic coverage-based rarefaction and extrapolation analysis. 

* `STEP4`. Evenness among species abundances. 


These are the compositions of `iNEXT.4steps`. Here we will also introduce functions for `STEP1` and `STEP4`, particularly. If you want to grasp the functions of `STEP2` or `STEP3` only, then please refer to the package [iNEXT.3D](https://cran.r-project.org/web/packages/iNEXT.3D/index.html) from CRAN for more details. `iNEXT.3D` features two statistical analyses (non-asymptotic and asymptotic):

1. A non-asymptotic approach based on interpolation and extrapolation for 3D diversity (i.e., Hill-Chao
numbers)

\code{iNEXT.3D} computes the estimated 3D diversity for standardized samples with a common sample size or sample completeness. This approach aims to compare diversity estimates for equally-large (with a common sample size) or equally-complete (with a common sample coverage) samples; it is based on the seamless rarefaction and
extrapolation (R/E) sampling curves of Hill-Chao numbers for q = 0, 1 and 2.

2. An asymptotic approach to infer asymptotic 3D diversity (i.e., Hill-Chao numbers)


`iNEXT.4steps` provides four functions for users to compute and visualize sample completeness curves and evenness curves: `Completeness` and `Evenness` for computation, and  `ggCompleteness` and `ggEvenness` for visualization. The most comprehensive function `iNEXT4steps` gathers `iNEXT3D`, `AO3D`, `Completeness` and `Evenness` into a summary table and a figure. In this document, we will give a quick introduction demonstrating how to run these functions. Detailed information about these function settings is provided in the `iNEXT.4steps` manual, which will be submitted to [CRAN]. See Chao et al. (2020) to get more details.


## How to cite
If you publish your work based on results from iNEXT.4steps package, you should make references to the following methodology paper and the package:

- Chao, A., Y. Kubota, D. Zeleny, C.-H. Chiu, C.-F. Li, B. Kusumoto, M. Yasuhara, S. Thorn, C.-L. Wei, M. J. Costello, and R. K. Colwell (2020). Quantifying sample completeness and comparing diversities among assemblages. Ecological Research, 35, 292-314.


## SOFTWARE NEEDED TO RUN INEXT IN R
- Required: [R](https://cran.r-project.org/)
- Suggested: [RStudio IDE](https://posit.co/products/open-source/rstudio/#Desktop)


## HOW TO RUN INEXT.4STEPS:
The `iNEXT.4steps` package will be submitted to [CRAN] or can be downloaded from Anne Chao's [iNEXT.4steps_github](https://github.com/AnneChao/iNEXT.4steps). For a first-time installation, some additional packages must be installed and loaded; see package manual.


```{r eval=FALSE}
## install iNEXT.4steps package from CRAN
# install.packages("iNEXT.4steps")  # Coming soon

## install the latest version from github
install.packages('devtools')
library(devtools)
install_github('AnneChao/iNEXT.4steps')

## import packages
library(iNEXT.4steps)
```

An online version of iNEXT.4steps (https://chao.shinyapps.io/iNEXT_4steps/) is also available for users without R background.


## SIMPLE STEPS BRIEF

- **Step 1: Sample completeness profile**

`Sample Completeness` represent the proportion of observed species in the population (Chao et al., 2020). Usually, the sampling data represents the abundant species in the population so that we will ignore the rare species. Here we will use Turing’s sample coverage theory to reconstruct the population proportion. Besides, sample completeness can correspond to order q, which is an weight index. When order q tends to zero, then we will give more weight to rare species. If order q tends to unity, then we will equally treat each species, which is also called sample coverage at unity. In contrast, if order q tends to larger than unity, we will give more weights to abundant species. By sample completeness, we can easily plot the estimated curve with respect to order q and associated 95% confidence interval.


- **Step 2.1 and step 3: Size-based and coverage-based Interpolation and Extrapolation** 

`Interpolation and Extrapolation (iNEXT)` focuses on three measures of order q: species richness (q = 0), Shannon diversity (q = 1, the exponential of Shannon entropy), and Simpson diversity (q = 2, the inverse of Simpson concentration) (Chao and Jost, 2012; Chao et al. 2014). For each diversity measures, `iNEXT` uses observed sample to compute expected diversity estimates and associated 95% confidence intervals according two different unit types of rarefaction and extrapolation (R/E):

1.	Sample-size-based R/E sampling curves versus diversity in each order q.

2.	Coverage-based R/E sampling curves versus diversity in each order q. 

A unified framework based on Hill numbers (for TD) and their generalizations (Hill-Chao numbers, for PD and
FD) is adopted to quantify 3D. In this framework, TD quantifies the effective number of species, PD quantifies the effective total branch length, mean-PD (PD divided by tree depth) quantifies the effective number of lineages, and FD quantifies the effective number of virtual functional groups (or functional “species”). Thus, TD, mean-PD, and FD are all in the same units of species/lineage equivalents and can be meaningfully compared. For more particular usage about `iNEXT.3D`, please refer to Chao et al. (2021).


- **Step 2.2: Asymptotic diversity profile** 

`Asymptotic Diversity (or called Hill numbers)` computes the estimated asymptotic 3D diversity and also plots 3D diversity profiles (q-profiles) for q between 0 and 2, in comparison with the observed diversity. Typically, the asymptotic estimates for q ≥ 1 are reliable, but for q < 1 (especially for q = 0, species richness), the asymptotic estimates represent only lower bounds. `Asymptotic Diversity` also features a time-profile (which depicts the observed and asymptotic estimate of PD or mean PD with respect to reference times), and a tau-profile (which depicts the observed and asymptotic estimate of FD with respect to threshold level tau).


- **Step 4: Evenness profile**

`Evenness` is an function to compute whether a assemblage is uniform or not. There are five main classes according to different transformation by species and diversity (Chao and Ricotta, 2019). In these five classes, they all have range from zero to one. When the value reaches to minimum (zero), it means that the assemblage tends to uneven. On the contrary, when the value reaches to maximum (one), it means that the assemblage tends to even. Evenness considers different order q under each classes. When order q tends to zero, we will focus on rare species. In other sides, when order q tends to far from zero, then we will give more weights on abundant species. For each sampling data, it usually has biase because of unobserved rare species. Here, we propose a "standardized sample coverage" (default is `Cmax`, which means the minimum sample coverage among all samples extrapolated to double reference sizes) as a judged criterion. According to this criterion, we can also plot the Evenness curves with order q and associated 95% confidence interval.


## <span style="color:red;">DATA INPUT FORMAT</span>

### Species abundance/incidence data format

For `iNEXT.3D` package, information on species identity (or any unique identification code) and assemblage affiliation is required for PD and FD. Two types of species abundance/incidence data are supported:

1. Individual-based abundance data (`datatype = "abundance"`): When there are multiple assemblages, in addition to the assemblage/site names (as column names) and the species names (as row names), species abundance data (reference sample) can be input as a species (in rows) by assemblage (in columns) matrix/data.frame or a list of species abundance vectors. In the special case that there is only one assemblage, all data should be read in one column.

2. Sampling-unit-based incidence data: There are two kinds of input data.  
  (a) Sampling-unit-based incidence data: Incidence-raw data (`datatype = "incidence_raw"`): for each assemblage, input data for a reference sample consist of a species-by-sampling-unit matrix, in addition to the sampling-unit names (as column names) and the species names (as row names). When there are N assemblages, input data consist of N lists of matrices, and each matrix is a species-by-sampling-unit matrix. Each element in the incidence raw matrix is 1 for a detection, and 0 for a non-detection. Input a matrix which combines data for all assemblages is allowed, but the argument `nT` in the function `iNEXT3D` must be specified so that the number of sampling units in each assemblage is specified.

  (b) Sampling-unit-based incidence-frequency data (`datatype = "incidence_freq"`): input data for each assemblage consist of species sample incidence frequencies (i.e., row sums of the corresponding incidence raw matrix). When there are N assemblages, input data consist of an (S+1) by N matrix, or N lists of species incidence frequencies. The first entry of each column/list must be the total number of sampling units, followed by the species incidence frequencies.

For example, the dataset `Brazil_abun_data` included in the `iNEXT.3D` package consists of species sample abundances of two assemblages/habitats: "Edge" and "Interior". Run the following code to view the first 15 rows of the abundance data. 

```{r eval=FALSE}
data("Brazil_abun_data")
Brazil_abun_data
```

```{r echo=FALSE}
data("Brazil_abun_data")
Brazil_abun_data[1:15,]
```


We use incidence frequency data (`Woody_plants`) collected from four sites, namely `"Upper_cloud"`, `"Lower_cloud"`, `"Lowland"`, and `"Monsoon"`, as an example. Note that the first row of incidence frequency data should be the total sampling units for each assemblage. Run the following code to view the first 16 rows and first 2 columns of the incidence frequency data. 


```{r eval=FALSE}
data("Woody_plants")
Woody_plants
```

```{r echo=FALSE}
data("Woody_plants")
Woody_plants[1:16, 1:2]
```


### Phylogenetic tree format for PD
To perform PD analysis, the phylogenetic tree (in Newick format) spanned by species observed in the pooled data is required. For the data `Brazil_abun_data`, the phylogenetic tree for all observed species (including species in both Edge and Interior habitats) is stored in the file `Brazil_phylo_tree`. A partial list of the tip labels and node labels are shown below. 

```{r}
data("Brazil_phylo_tree")
Brazil_phylo_tree
```


### Species pairwise distance matrix format for FD
To perform FD analysis, the species-pairwise distance matrix (Gower distance computed from species traits) for species observed in the pooled data is required in a matrix/data.frame format. For the data `Brazil_abun_data`, the distance matrix for all species (including species in both Edge and Interior habitats) is stored in the file `Brazil_rainforest_dist_matrix`. The distance matrix for the first 3 Brazil rainforest tree species is shown below. 


```{r eval=FALSE}
data("Brazil_distance_matrix")
Brazil_distance_matrix
```

```{r echo=FALSE}
data("Brazil_distance_matrix")
round(Brazil_distance_matrix[1:3,1:3], 3)
```


## <span style="color:red;">MAIN FUNCTION iNEXT4steps()</span>

We first describe the main function `iNEXT4steps()` with default arguments: 

```{r eval=FALSE}
iNEXT4steps(data, diversity = "TD", q = seq(0, 2, 0.2), datatype = "abundance", 
            nboot = 50, conf = 0.95, nT = NULL, PDtree = NULL, PDreftime = NULL, PDtype = 'meanPD', 
            FDdistM = NULL, FDtype = 'AUC', FDtau = NULL, FDcut_number = 50, details = FALSE)
```

The arguments of this function are briefly described below, and will explain details by illustrative examples in later text. 
<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For `datatype = "abundance"`, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For `datatype = "incidence_freq"`, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For `datatype = "incidence_raw"`, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. 
</td>
</tr>
<tr class="even">
<td align="center"><code>diversity</code></td>
<td align="left">selection of diversity type: `TD` = Taxonomic diversity, `PD` = Phylogenetic diversity, and `FD` = Functional diversity.</td>
</tr>
<tr class="odd">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders for q-profile output. Default is <code>seq(0, 2, by = 0.2)</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (`datatype = "abundance"`), sampling-unit-based incidence frequencies data (`datatype = "incidence_freq"`), or species by sampling-units incidence matrix (`datatype = "incidence_raw"`) with all entries being 0 (non-detection) or 1 (detection)</td>
</tr>
<tr class="odd">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="even">
<td align="center"><code>conf</code></td>
<td align="left">a positive number < 1 specifying the level of confidence interval. Default is 0.95.</td>
</tr>
<tr class="odd">
<td align="center"><code>nT</code></td>
<td align="left">(required only when <code>datatype = "incidence_raw"</code> and input data in a single matrix/data.frame) a vector of positive integers specifying the number of sampling units in each assemblage. If assemblage names are not specified (i.e., <code>names(nT) = NULL</code>), then assemblages are automatically named as "assemblage1", "assemblage2",..., etc.</td>
</tr>
<tr class="even">
<td align="center"><code>PDtree</code></td>
<td align="left">(required argument for <code>diversity = "PD"</code>), a phylogenetic tree in Newick format for all observed species in the pooled assemblage.</td>
</tr>
<tr class="odd">
<td align="center"><code>PDreftime</code></td>
<td align="left">(argument only for <code>diversity = "PD"</code>), a vector of numerical values specifying reference times for PD. Default is <code>NULL</code> (i.e., the age of the root of <code>PDtree</code>).</td>
</tr>
<tr class="even">
<td align="center"><code>PDtype</code></td>
<td align="left">(argument only for <code>diversity = "PD"</code>), select PD type: <code>PDtype = "PD"</code> (effective total branch length) or <code>PDtype = "meanPD"</code> (effective number of equally divergent lineages). Default is <code>"meanPD"</code>, where <code>meanPD = PD/tree depth</code>.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDdistM</code></td>
<td align="left">(required argument for <code>diversity = "FD"</code>), a species pairwise distance matrix for all species in the pooled assemblage.</td>
</tr>
<tr class="even">
<td align="center"><code>FDtype</code></td>
<td align="left">(argument only for <code>diversity = "FD"</code>), select FD type: <code>FDtype = "tau_values"</code> for FD under specified threshold values, or <code>FDtype = "AUC"</code> (area under the curve of tau-profile) for an overall FD which integrates all threshold values between zero and one. Default is <code>"AUC"</code>.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDtau</code></td>
<td align="left">(argument only for <code>diversity = "FD"</code> and <code>FDtype = "tau_values"</code>), a numerical vector between 0 and 1 specifying tau values (threshold levels). If <code>NULL</code> (default), then threshold is set to be the mean distance between any two individuals randomly selected from the pooled assemblage (i.e., quadratic entropy).</td>
</tr>
<tr class="even">
<td align="center"><code>FDcut_number</code></td>
<td align="left">(argument only for <code>diversity = "FD"</code> and <code>FDtype = "AUC"</code>), a numeric number to cut [0, 1] interval into equal-spaced sub-intervals to obtain the AUC value by integrating the tau-profile. Equivalently, the number of tau values that will be considered to compute the integrated AUC value. Default is <code>FDcut_number = 50</code>. A larger value can be set to obtain more accurate AUC value.</td>
</tr>
<tr class="odd">
<td align="center"><code>details</code></td>
<td align="left">a logical variable to decide whether do you want to print out the detailed value for each plots, default is `FALSE`.</td>
</tr>

</tbody>
</table>


Data type has following formats, such as a vector gathered by factors (`abundance` and `incidence_freq`), a matrix/data frame with species versus a assemblage (`abundance` `incidence_freq`, and `incidence_raw`), a list of several vectors (`abundance` and `incidence_freq`), or a list correspond to a assemblage (`incidence_raw`). `data` should  comform the format of each datatype. When `datatype = "incidence_raw"` and class of data is matrix/data frame, user should input `nT` for each assemblage which represents sampling units.

`diversity`  contains three attributes diversity dimensions: `Taxonomic diversity`, `Phylogenetic diversity`, and `Functional diversity`. User should choose one `diversity`: `TD` means `Taxonomic diversity`, `PD` means `Phylogenetic diversity` under a specified reference time (default is root height), and `FD` means `Functional diversity`. For the `"Functional diversity"`, `FDtype = "AUC"` consider overall functional diversity which integrates all threshold values between 0 to 1. And `FDtype = "tau_values"` computes functional diversity under specified thresholds.

When `diversity = "PD"`, user should input `PDtree` Newick format data for all observed species. When `diversity = "FD"`, user should input `FDdistM` data matrix. Each element of the matrix is the pairwise distance between any two observed species. And the species identification names should be listed as row names and column names of the matrix.

`nboot` is applied to get confidence interval, which is estimated by bootstrap method. `details` means a logical setting whether print out the computation value of all figures.

The output of `iNEXT4steps` will have three parts (if `details = TRUE`): `$summary`, `$figure`, and `$details`. It may take some time to compute when data size is large or `nboot` is large.


## <span style="color:blue;">TAXONOMIC DIVERSITY (TD): 4 STEPS VIA EXAMPLES</span>

`"abundance Data"` is used for a random sampling scheme. If the species has aggregation effect, such as trees or plants, then set `datatype = "incidence_freq"` or `datatype = "incidence_raw"`. First, we use data `Spider` to compute taxonomic diversity.

### EXAMPLE 1: TD 4 steps for abundance data 

Datasets `Spider` were sampled in a mountain forest ecosystem in the Bavarian Forest National Park, Germany (Thorn et al. 2016, 2017).
A total of 12 experimental plots were established in "closed forest" stands (6 plots) and "open forest" stands with naturally occurring gaps and edges (6 plots) to assess the effects of microclimate on communities of epigeal (ground-dwelling) spiders. Epigeal spiders were sampled over three years with four pitfall traps in each plot, yielding a total of 3171 individuals belonging to 85 species recorded in the pooled habitat. More details refer to data Source : A mountain forest ecosystem in the Bavarian Forest National Park, Germany (Thorn et al. 2016, 2017).

Run the following code to get the output from iNEXT4steps.

```{r align="center",out.width="100%"}
data(Spider)
out1 <- iNEXT4steps(data = Spider, diversity = "TD", datatype = "abundance")
out1$summary
out1$figure[[6]]
```

`$summary` lists all biological summaries according to Chao et al. (2020). There are four parts corresponding to each step in the paper. They analysis and explain biological data from different and superimposed side. User can easily compare difference between each assemblages. 

`$figure` visualize the statistics by continuous curves. From the above five figures, `iNEXT4stpes` provides a standard analysis process from figure (a) to figure (e). User can analyze the process of biodiversity through these figures. 

`$details` contains four parts: `Sample Completeness`, `iNEXT`, `Asymptotic Diversity`, `Evenness`. They are the computing values which are used to plot each figure in `$figure`. 


### EXAMPLE 2: TD 4 steps for incidence data 

Incidence raw data is a species-by-incidence-sampling-units matrix/data.frame. We split a space into several quadrats and only record whether the species is detected or undetected in each quadrat. According to this sampling scheme, `incidence_raw` data has only value "zero" (undetected) or "one" (detected) in matrix/data frame (species by assemblages). `incidence_freq` data is the total incidence frequency for each species (i.e., row sums of the corresponding incidence raw matrix). `incidence_freq` data should contain total sampling units (number of quadrats) in the first row/entry.

*Note:* The phylogenetic diversity can only accept `datatype = "incidence_raw"` for incidence-based data.

Datasets `Woody plants` are a subset of The National Vegetation Database of Taiwan (AS-TW-001), sampled between 2003 and 2007 within the first national vegetation inventory project (Chiou et al. 2009). Over 3600 vegetation plots, each 20x20-m in area, were set up in various locations in Taiwan, and all woody plant individuals taller than 2 meters were recorded in each plot. For illustration here, we selected only plots belonging to two vegetation types (according to Li et al. 2013): Pyrenaria-Machilus subtropical winter monsoon forest and Chamaecyparis montane mixed cloud forest, sampled in the northern part of Taiwan (in ecoregions 7 and 8 according to Su 1985).

Run the following code to get the output from iNEXT4steps.

```{r align="center", out.width="100%"}
data(Woody_plants)
out2 <- iNEXT4steps(data = Woody_plants[,c(1,4)], diversity = "TD", datatype = "incidence_freq")
out2$summary
out2$figure[[6]]
```


## <span style="color:blue;">PHYLOGENETIC DIVERSITY (PD): 4 STEPS VIA EXAMPLES</span>

Here use abundance data: `"Brazil_abun_data"` as example. `"Brazil_abun_data"` data has two main communities: "Edge", "Interior". Here we also provide phylogenetic tree (`Brazil_phylo_tree`) data for `"Brazil_abun_data"` data to compute phylogenetic diversity. 

Run the following code to get the output from iNEXT4steps.

```{r align="center", out.width="100%"}
data(Brazil_abun_data)
data(Brazil_phylo_tree)
out3 <- iNEXT4steps(data = Brazil_abun_data, diversity = "PD", datatype = "abundance", nboot = 20, 
                    PDtree = Brazil_phylo_tree)
out3$summary
out3$figure[[6]]
```


## <span style="color:blue;">FUNCTIONAL DIVERSITY (FD): 4 STEPS VIA EXAMPLES</span>

Here use abundance data `Brazil_abun_data` and its pairwise distance matrix (`Brazil_distance_matrix`) to compute functional diversity. Under `FDtype = "tau_values"`, user can key in `FDtau` as thresholds (default is `Cmax`). Under `FDtype = "AUC"`, it considers overall functional diversity which integrates all threshold values between 0 to 1. And `FDtype = "tau_values"` computes functional diversity under specified thresholds.

Run the following code to get the output from iNEXT4steps.

```{r align="center", out.width="100%"}
data(Brazil_abun_data)
data(Brazil_distance_matrix)
out4 <- iNEXT4steps(data = Brazil_abun_data, diversity = "FD", datatype = "abundance", nboot = 20, 
                    FDdistM = Brazil_distance_matrix, FDtype = 'tau_values')
out4$summary
out4$figure[[6]]
```


## <span style="color:red;">FUNCTION Completeness: SAMPLE COMPLETENESS PROFILES</span>

`iNEXT.4steps` provides function `Completeness()` to compute estimated sample completeness with order q. The arguments is below: 

```{r eval=FALSE}
Completeness(data, q = seq(0, 2, 0.2), datatype = "abundance", nboot = 50, conf = 0.95, nT = NULL)
```

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For `datatype = "abundance"`, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For `datatype = "incidence_freq"`, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For `datatype = "incidence_raw"`, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. 
</td>
</tr>
<tr class="even">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders. Default is <code>seq(0, 2, by = 0.2)</code>.</td>
</tr>
<tr class="odd">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (`datatype = "abundance"`), sampling-unit-based incidence frequencies data (`datatype = "incidence_freq"`), or species by sampling-units incidence matrix (`datatype = "incidence_raw"`) with all entries being 0 (non-detection) or 1 (detection)</td>
</tr>
<tr class="even">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="odd">
<td align="center"><code>conf</code></td>
<td align="left">a positive number < 1 specifying the level of confidence interval. Default is 0.95.</td>
</tr>
<tr class="even">
<td align="center"><code>nT</code></td>
<td align="left">(required only when `datatype = "incidence_raw"` and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc.</td>
</tr>

</tbody>
</table>

## <span style="color:red;">FUNCTION ggCompleteness(): GRAPHIC DISPLAYS OF SAMPLE COMPLETENESS PROFILES</span>

`iNEXT.4steps` also provides a visualized function `ggCompleteness` to plot the output from `Completeness()`:

```{r eval=FALSE}
ggCompleteness(output)
```

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>output</code></td>
<td align="left">the output of the function <code>Completeness</code>.
</td>

</tbody>
</table>


Following are two simple examples for functions `Completeness` and `ggCompleteness`.  

### Sample completeness profiles for abundance data

Use abundance data `Spider` to compute estimated sample completeness.

```{r}
data(Spider)
out1 <- Completeness(data = Spider, datatype = "abundance")
out1
```

The following commands plot sample completeness curves, along with its confidence interval for q between 0 to 2.

```{r out.width="70%",fig.height=8}
ggCompleteness(out1)
```


### Sample completeness profiles for incidence data

Use incidence frequency data `woody plants` to compute sample completeness.

```{r}
data(Woody_plants)
out2 <- Completeness(data = Woody_plants[,c(1,4)], datatype = "incidence_freq")
out2
```

The following commands plot sample completeness curves, along with its confidence interval for q between 0 to 2.

```{r out.width="70%",fig.height=8}
ggCompleteness(out2)
```



## <span style="color:red;">FUNCTION Evenness: EVENNESS PROFILES</span>

`iNEXT.4steps` provides the function `Evenness()` to compute observed eveness or estimated evenness under specified sample coverage. The argument is below: 

```{r eval=FALSE}
Evenness(data, q = seq(0, 2, 0.2), datatype = "abundance", method = "Estimated",
         nboot = 50, conf = 0.95, nT = NULL, E.class = 1:5, SC = NULL)
```

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For `datatype = "abundance"`, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For `datatype = "incidence_freq"`, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For `datatype = "incidence_raw"`, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. 
</td>
</tr>
<tr class="even">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders. Default is <code>(0, 0.2, 0.4,...,2)</code>.</td>
</tr>
<tr class="odd">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (`datatype = "abundance"`), sampling-unit-based incidence frequencies data (`datatype = "incidence_freq"`), or species by sampling-units incidence matrix (`datatype = "incidence_raw"`) with all entries being 0 (non-detection) or 1 (detection)</td>
</tr>
<tr class="even">
<td align="center"><code>method</code></td>
<td align="left">a binary calculation method with <code>"Estimated"</code> or <code>"Observed"</code>.</td>
</tr>
<tr class="odd">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is <code>50</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>conf</code></td>
<td align="left">a positive number < <code>1</code> specifying the level of confidence interval. Default is <code>0.95</code>.</td>
</tr>
<tr class="odd">
<td align="center"><code>nT</code></td>
<td align="left">(required only when `datatype = "incidence_raw"` and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc.</td>
</tr>
<tr class="even">
<td align="center"><code>E.class</code></td>
<td align="left">an integer vector between 1 to 5.</td>
</tr>
<tr class="odd">
<td align="center"><code>SC</code></td>
<td align="left">(required only when method = "Estimated") a standardized coverage for calculating estimated evenness. If <code>NULL</code>, then this function computes the diversity estimates for the minimum sample coverage among all samples extrapolated to double reference sizes (Cmax).</td>
</tr>

</tbody>
</table>


## <span style="color:red;">FUNCTION ggEvenness(): GRAPHIC DISPLAYS OF EVENNESS PROFILES</span>

`iNEXT.4steps` provide a function `ggEvenness()` to plot the output from `Evenness()`.

```{r eval=FALSE}
ggEvenness(output)
```

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>output</code></td>
<td align="left">the output of the function <code>Evenness</code>.
</td>

</tbody>
</table>

Following are two simple examples for functions `Evenness` and `ggEvenness`. 

### Evenness profiles for abundance data

Use abundance data `Spider` to compute estimated evenness under `C = Cmax`. Here only show the output for first class of evenness.

```{r}
data(Spider)
out1 <- Evenness(data = Spider, datatype = "abundance", 
                 method = "Estimated", SC = NULL, E.class = 1:5)
out1$E1
```

The following commands plot the evenness curves, along with its confidence interval for q between 0 to 2.

```{r out.width="100%",fig.height=8}
ggEvenness(out1)
```



### Evenness profiles for incidence data

Use incidence frequency data `Woody plants` to calculate estimated evenness under `C = NULL` (Cmax). Here only show the output for first class of evenness.


```{r}
data(Woody_plants)
out2 <- Evenness(data = Woody_plants[,c(1,4)], datatype = "incidence_freq", 
                 method = "Estimated", SC = NULL, E.class = 1:5)
out2$E1
```

The following commands plot the evenness curves, along with its confidence interval for q between 0 to 2.

```{r out.width="100%",fig.height=8}
ggEvenness(out2)
```



## License
The iNEXT.4steps package is licensed under the GPLv3. To help refine `iNEXT.4steps`, your comments or feedback would be welcome (please send them to Anne Chao or report an issue on the iNEXT.4steps github [iNEXT.4steps_github](https://github.com/AnneChao/iNEXT.4steps). 


## References
- Chao, A., Gotelli, N. G., Hsieh, T. C., Sander, E. L., Ma, K. H., Colwell, R. K. and Ellison, A. M. (2014). Rarefaction and extrapolation with Hill numbers: a framework for sampling and
estimation in species biodiversity studies. Ecological Monographs 84, 45-67. 

- Chao, A., Henderson, P. A., Chiu, C.-H., Moyes, F., Hu, K.-H., Dornelas, M and Magurran, A. E. (2021). Measuring temporal change in alpha diversity: a framework integrating taxonomic, phylogenetic and functional diversity and the iNEXT.3D standardization. Methods in Ecology and Evolution, 12, 1926-1940.

- Chao, A. and Jost. L. (2012) Coverage-based rarefaction and extrapolation: standardizing samples by completeness rather than size. Ecology, 93, 2533-2547.

- Chao, A. and Jost, L. (2015). Estimating diversity and entropy profiles via discovery rates of new species. Methods in Ecology and Evolution, 6, 873-882.

- Chao, A. and Ricotta, C. (2019). Quantifying evenness and linking it to diversity, beta diversity, and similarity. Ecology, 100(12), e02852.

- Chao, A., Y. Kubota, D. Zeleny, C.-H. Chiu, C.-F. Li, B. Kusumoto, M. Yasuhara, S. Thorn, C.-L. Wei, M. J. Costello, and R. K. Colwell (2020). Quantifying sample completeness and comparing diversities among assemblages. Ecological Research, 35, 292-314.

- Chiou, C.-R., Hsieh, C.-F., Wang, J.-C., Chen, M.-Y., Liu, H.-Y., Yeh, C.-L., ... Song, M. G.-Z. (2009). The first national vegetation inventory in Taiwan. Taiwan Journal of Forest Science, 24, 295–302.

- Li, C.-F., Chytry, M., Zeleny, D., Chen, M. -Y., Chen, T.-Y., Chiou, C.-R., ... Hsieh, C.-F. (2013). Classification of Taiwan forest vegetation. Applied Vegetation Science, 16, 698–719. https://doi.org/10.1111/avsc.12025

- Su, H. -J. (1985). Studies on the climate and vegetation types of the natural forests in Taiwan (III) A Scheme of Geographical Climatic Regions. Quarterly Journal of Chinese Forestry, 18, 33–44.

- Thorn, S., Bassler, C., Svoboda, M., & Müller, J. (2017). Effects of natural disturbances and salvage logging on biodiversity - lessons from the bohemian Forest. Forest Ecology and Management, 388, 113–119. https://doi.org/10.1016/j.foreco.2016.06.006

- Thorn, S., BuBler, H., Fritze, M. -A., Goeder, P., Muller, J., WeiB, I., & Seibold, S. (2016). Canopy closure determines arthropod assemblages in microhabitats created by windstorms and salvage logging. Forest Ecology and Managemen


